---
title: "prepare_data"
author: "Aimara Planillo"
date: "2023-06-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set workspace}
source("./R/source_packages.R")

procdata_wd <- file.path("./data_proc")
```


# Prepare wolf data for demographic analysis

In all cases, we are going to use the monitoring or wolf year: 1st May - 31st April next year
For example, year 2020 refers to teh time between 1st May 2020 and 31st April 2021


# Prepare Survival Data

We are going to prepare the data for the analysis. For this, 
we filter the unusable records and add the neccesary information.
In summary, final dataset will have:
- All individuals born before 2021 and with known birth date
- Only individuals in Germany (due to explantory variables)
- We use standard birth date: 1. May
- We use SDat2 as last date: death for dead ind, and last date for censored ind
- We assume death within a week of last date
- Prepare the response variable
- Add the explanatory variables: habitat suitability, territory density, sex, age, season


**RESPONSE variable**
weeks_date: 
- Number of weeks since birth until death of last record of the individual

**EXPLANATORY variables**
_Status_: 
- Dead = 1
- Alive = Censored = 0

_Sex_: 
- male  
- female 

_Season_:
- Summer = May-Oct (when new pups are born and stay with pack)
- Winter = Nov-Apr 

_Characteristics of Natal territory_: coord_x_first, coord_y_first
- Definition of natal territory: 
    - Coordinates of birth if exist
    - First Known location when no birth coordinates are provided
- _Natal territory HS_: Average habitat suitability values in buffer of 8km (represents territory)
- _Natal territory density_: Territory density in buffer of 50 km in the birth year

_Characteristics of Final territory_: coord_x_last, coord_y_last
- Definition of final territory: 
    - Coordinates of death if exist
    - Last known location when no death coordinates are provided
- _Final territory HS_: Average habitat suitability values in buffer of 8km (represents territory)
- _Final territory density_:Territory density in buffer of 50 km in the death or last record year


## Load and prepare data frame
```{r}
rawsurv <- read.csv("./data_raw/data_survival_rawtable.csv")

# set unclear Sex cases to NA
surv_sex_tmp <- rawsurv %>% 
  mutate(Sex = ifelse(Sex == "m?" | Sex == "f?", NA, Sex))

# remove individuals that were born in 2021 or later and undefined Sex
surv_2021_tmp <- surv_sex_tmp %>%
  filter(!is.na(Sex)) %>% 
  filter(ju_end < 2021) 

table(surv_2021_tmp$Sex)

# remove individuals with unknown year of birth
surv_2021_birth <- surv_2021_tmp %>% 
  filter(ju_start == ju_end)

## get coordinates of wolves: birth place is known. If unknown, first known location
surv_coord <- surv_2021_birth %>%
  mutate(coord_x_first = case_when(
    birth_x != "not.found" ~ as.numeric(birth_x), 
    birth_x == "not.found" ~ first_x), 
    coord_y_first = case_when(
    birth_y != "not.found" ~ as.numeric(birth_y), 
    birth_y == "not.found" ~ first_y)) %>% 
  mutate(coord_x_last = last_x, 
         coord_y_last = last_y)

## check for missing coordinates
surv_coord %>% 
  filter(is.na(coord_x_first)) # these are individuals with origin outside Germany

surv_coord %>% 
  filter(is.na(coord_x_last)) # these are individuals with only one sighting

nrow(surv_coord) # 1098

head(surv_coord)

## add bird date, death date and number of weeks of data
surv_dates_tmp <- surv_coord %>% 
  ## add dates
  mutate(birth_date_char = paste0("01.05.", ju_end),
         SDat2_char = as.character(SDat2),
         birth_date = as.Date(birth_date_char, format = "%d.%m.%Y"), 
         last_date = as.Date(SDat2_char, format = "%Y%m%d")) %>% 
  ## add month of death (for season)
  mutate(death_month = month(last_date)) %>% 
  ## add years of birth and death
  # they are born at the beginnig of hte wolf year but
  # Because we are using monitoring years, we need to transform the death year to monitoring year 
  mutate(birth_year = year(birth_date), 
         death_year = case_when(
    month(last_date) < 5 ~ year(last_date)-1,
    month(last_date) > 4 ~ year(last_date))) %>%
  ## add response variable: weeks alive (weeks_date)
  mutate(weeks_date = interval(birth_date, last_date) %/% weeks(1)) %>% 
  # add status first base on Mortality on april 2021
  mutate(status = case_when(
    Mortality_2020 == "alive" ~ 0, 
    Mortality_2020 == "verified" ~ 1, 
    Mortality_2020 == "unverified" ~ 0)
  ) %>% 
  mutate(season = case_when(
    death_month %in% c(5:10) ~ "Summer",
    death_month %in% c(1:4, 11,12) ~ "Winter"
    )) %>% 
  ## keep only relevant variables
    dplyr::select(Individual, Pack_origin, weeks_date, status, 
                coord_x_first, coord_y_first, coord_x_last, coord_y_last, 
                Sex, birth_date, birth_year, last_date, death_year, death_month, 
                SDat1, SDat2, MJ1, MJ2, season, region_origin)

head(surv_dates_tmp)
nrow(surv_dates_tmp) # 1098

surv_dates_tmp %>% 
  filter(!is.na(coord_x_last))

# add last location = first location for individuals with only one data
surv_dates_tmp2 <- surv_dates_tmp %>% 
  mutate(coord_x_last = case_when(
    is.na(coord_x_last) & MJ1 == MJ2 ~ coord_x_first,
    TRUE ~ coord_x_last)) %>% 
  mutate(coord_y_last = case_when(
    is.na(coord_y_last) & MJ1 == MJ2 ~ coord_y_first,
    TRUE ~ coord_y_last))


# Remove individuals with negative weeks, as this is a mistake in the data
surv_dates_tmp2 %>% 
  filter(weeks_date < 0)

nrow(surv_dates_tmp2) # 1098
str(surv_dates_tmp2)

## remove unknown coordinates and ind from outside germany
unique(surv_dates_tmp2$region_origin)

surv_dates_tmp3 <- surv_dates_tmp2 %>% 
  filter(!is.na(coord_x_first)) %>%
  filter(!region_origin %in% c("EU_DK", "EU_NW", "EU_ME"))

nrow(surv_dates_tmp3) # [1] 1061
unique(surv_dates_tmp3$region_origin)

surv_dates_tmp3 %>% 
  group_by(region_origin) %>% 
  summarise(n = n())

## assign 2020 as last effective year
surv_dates_tmp3 <- surv_dates_tmp3 %>% 
  mutate(eff_last_year = case_when(
    death_year < 2021 ~ as.numeric(death_year),
    death_year > 2020 ~ 2020))
```

## SPATIAL DATA

### Divide data into first and last locations and make spatial

```{r Check that all individuals are in Germany}
## load german border
german_states <- read_sf("./data_raw/sf_german_regions_3035.gpkg", crs = 3035)

## make coordinates spatial objects
first_sf <- st_as_sf(surv_dates_tmp3, coords = c("coord_x_first", "coord_y_first"), crs = 4326) %>% 
  st_transform(crs = 3035)
last_sf <- st_as_sf(surv_dates_tmp3, coords = c("coord_x_last", "coord_y_last"), crs = 4326) %>% 
  st_transform(crs = 3035)

last_sf %>% filter(Individual == "GW1905f")
  
tmap_mode("view")
tm_shape(german_states) +
  tm_borders() +
tm_shape(first_sf) +
  tm_dots("red", size = 0.2) + 
  tm_shape(last_sf) +
  tm_dots(size = 0.1)

## We see some data outside the German borders
```

### Remove individuals outside Germany
```{r remove outside german locations}
## extract regions for first locations
out_first <- st_intersection(first_sf, german_states)
out_first

out_first %>% 
    filter(Individual == "GW1267m")
## assign region of death to individuals
out_last_tmp <- st_intersection(last_sf, german_states) %>% 
  dplyr::select(Individual, name) %>% 
  st_drop_geometry()
out_last_tmp

out_last <- last_sf %>% 
  left_join(out_last_tmp, by = "Individual") %>% 
  mutate(region_death = name)

nrow(out_first) # [1] 1061
nrow(out_last) # [1] 1061

tm_shape(german_states) +
  tm_borders() +
tm_shape(out_first) +
  tm_dots("red", size = 0.2) + 
  tm_shape(out_last) +
  tm_dots(size = 0.1)
```


### Add Habitat suitability data 

```{r add habitat suitability values}
## load habitat suitability map (values*1000)
hs_map <- rast("./data_raw/ensemble_samebgpres_weightedmean_projection_3035.tif")
hs_map
plot(hs_map)

## see points of first and last location of individuals
tmap_mode("view")
tm_shape(hs_map) +
  tm_raster(palette = "viridis", alpha = 0.2) + 
  tm_shape(german_states) +
  tm_fill("name", alpha = 0.2) +
  tm_shape(out_first) +
  tm_dots("green") +
  tm_shape(out_last) +
  tm_dots("black")

## extract habitat suitability values in radius of 8 km of wolf first location
buffer_first <- st_buffer(out_first, dist = 8000) %>% 
  st_transform(3035)

buffer_first_hs <- exactextractr::exact_extract(hs_map, buffer_first, fun = "mean")

nrow(out_first) # 1061
length(buffer_first_hs) # 1061

first_hs <- out_first %>% 
  mutate(hs_8km_natal = buffer_first_hs / 1000)   # we set the hs values from 0 to 1 (original scale)

## extract habitat suitability values in radius of 8 km of wolf last location
buffer_last <- st_buffer(out_last, dist = 8000) %>% 
  st_transform(3035)

buffer_last_hs <- exactextractr::exact_extract(hs_map, buffer_last, fun = "mean")

nrow(out_last) # 1061
length(buffer_last_hs)  # 1061

last_hs <- out_last %>% 
  mutate(hs_8km_final = buffer_last_hs / 1000)  # we set the hs values from 0 to 1 (original scale)

## remove points with no data
last_hs <- last_hs %>% 
  filter(!is.na(hs_8km_final))
nrow(last_hs) # 1054
```

### Add wolf territory density in 50km radius of focal territory

```{r load territory data}
## load german border to cut buffers
german_border <- st_read("./data_raw/germany_3035.gpkg", crs = 3035)
plot(german_border)

## get territory data and make spatial data
wolf_terr_df <- read.csv("./data_raw/20220719_Territorien DBBW.csv")  #status 1-pack, 2-pair, 3-ind

## make wolf territory spatial, and remove single territories 
wolf_terr_sf <- wolf_terr_df %>% 
  # remove data after year 2020
  filter(Mjahr < 2021) %>% 
  # remove single territorial individuals
  filter(Status != 3) %>% 
  # make spatial object
  st_as_sf(coords = c("DECLNG", "DECLAT"), crs = 4326) %>% 
  # transform to projected coordinates
  st_transform(crs = 3035)

head(wolf_terr_df)
```

```{r get wolf density values in radius of first territory}
first_hs_3035 <- st_transform(first_hs, 3035)

## create a loop to extract the territory info for each year of first location
nyears_first <- length(unique(first_hs_3035$birth_year))

total_area_buffer50 <- pi*50^2
all_years_first_df <- data.frame()

for (i in 1:nyears_first){
  # select the year
  tmp_year = unique(first_hs_3035$birth_year)[i]
  
  # get individual 50km buffers in year
  poly50_tmp <- first_hs_3035 %>% 
    filter(birth_year == tmp_year) %>% 
    st_buffer(dist = 50000) 
  
  # remove the area of the buffers outside germany
  poly50_tmp_crop <- st_intersection(poly50_tmp, german_border)
  
  # add real area for each buffer in km2
  poly50_tmp_crop <- poly50_tmp_crop %>% 
    mutate(area_km2 = as.numeric(
      st_area(poly50_tmp_crop) %>% 
        set_units(km^2)))
  
  # get territory (packs and pairs) polygons in year
  wolf_terr_tmp <- wolf_terr_sf %>% 
    filter(Mjahr == tmp_year) %>% 
    st_buffer(dist = 8000)
  
  # get number of territory polygons intersecting the individual buffer
  intersect50_tmp <- st_intersects(poly50_tmp_crop, wolf_terr_tmp)
  n_int50_tmp <- lengths(intersect50_tmp)
  
  # make into a dataframe
  my_df_tmp <- cbind.data.frame(Individual = poly50_tmp$Individual,
                          nterr_buffer50_first = n_int50_tmp, 
                          area_buffer50_first = poly50_tmp_crop$area_km2,
                          Myear_first = tmp_year) 
  # get territory density using buffer size
  my_df_tmp_dens <- my_df_tmp %>% 
    mutate(nterr_dens_buffer50_first = nterr_buffer50_first / (area_buffer50_first) * total_area_buffer50) # Nterritories in full buffer area
  
  all_years_first_df <- rbind(all_years_first_df, my_df_tmp_dens)
  }


head(all_years_first_df)
nrow(all_years_first_df) #1061
nrow(first_hs_3035) #1061

unique(first_hs_3035$birth_year)
unique(all_years_first_df$Myear_first)

## joint n terr to data
first_hs_nterr <- first_hs_3035 %>% 
  left_join(all_years_first_df, by = c("Individual", "birth_year" = "Myear_first"))

nrow(first_hs_nterr) #1061
```


```{r get wolf density values in radius of last territory}
last_hs_3035 <- st_transform(last_hs, 3035)

nyears_last <- length(unique(last_hs_3035$eff_last_year ))

total_area_buffer50 <- pi*50^2

all_years_last_df <- data.frame()

for (i in 1:nyears_last){
  # select the year
  tmp_year = unique(last_hs_3035$eff_last_year)[i]
  
  # get individual 50km buffers in year
  poly50_tmp <- last_hs_3035 %>% 
    filter(eff_last_year == tmp_year) %>% 
    st_buffer(dist = 50000)
  
  # remove the area of the buffers outside germany
  poly50_tmp_crop <- st_intersection(poly50_tmp, german_border)
  
  # add real area for each buffer in km2
  poly50_tmp_crop <- poly50_tmp_crop %>% 
    mutate(area_km2 = as.numeric(
      st_area(poly50_tmp_crop) %>% 
        set_units(km^2)))
  
  # get territory (packs and pairs) polygons in year
  wolf_terr_tmp <- wolf_terr_sf %>% 
    filter(Mjahr == tmp_year) %>% 
    st_buffer(dist = 8000)
  
  # get number of territory polygons intersecting the individual buffer
  intersect50_tmp <- st_intersects(poly50_tmp_crop, wolf_terr_tmp)
  n_int50_tmp <- lengths(intersect50_tmp)
  
  # make into a dataframe
  my_df_tmp <- cbind.data.frame(Individual = poly50_tmp$Individual,
                          nterr_buffer50_last = n_int50_tmp, 
                          area_buffer50_last = poly50_tmp_crop$area_km2,
                          Myear_last = tmp_year) 
  # get territory density using buffer size
  my_df_tmp_dens <- my_df_tmp %>% 
    mutate(nterr_dens_buffer50_last = nterr_buffer50_last / (area_buffer50_last) * total_area_buffer50) #n territories in 100km2
           
  all_years_last_df <- rbind(all_years_last_df, my_df_tmp_dens)
  }

all_years_last_df
nrow(all_years_last_df) #1054
nrow(last_hs_3035) #1054

unique(last_hs_3035$death_year)
unique(all_years_last_df$Myear_last)

## joint n terr to data
last_hs_nterr <- last_hs_3035 %>% 
  left_join(all_years_last_df, by = c("Individual", "eff_last_year" = "Myear_last"))
```

```{r put together and save data}
## get individual ID
ind_first <- unique(first_hs_nterr$Individual)
ind_last <- unique(last_hs_nterr$Individual)

## retain only individuals in both datasets
ind_tokeep <- ind_first[which(ind_first %in% ind_last)]
length(ind_tokeep) # 1053

tmp_first <- first_hs_nterr %>% 
  filter(Individual %in% ind_tokeep) %>% 
  st_drop_geometry()

tmp_last <- last_hs_nterr %>% 
  filter(Individual %in% ind_tokeep) %>% 
  st_drop_geometry()

## merge datasets

surv_data <- tmp_first %>% 
  left_join(tmp_last %>% dplyr::select(Individual, coord_x_first, coord_y_first, 
                                       hs_8km_final, nterr_buffer50_last, area_buffer50_last, 
                                       nterr_dens_buffer50_last), 
            by = "Individual")
head(surv_data)
nrow(surv_data) # 1054
```

```{r save data}
# save data for later use
# write.csv(surv_data, paste0(procdata_wd, "/data_wolf_survival_hs_nterr.csv"),
#                  row.names = FALSE)
# surv_data <- read.csv(paste0(procdata_wd, "/data_wolf_survival_hs_nterr.csv"))
```

```{r remove sensitive info}
surv_data_git <- surv_data %>% 
  mutate(IndID = seq(1, nrow(surv_data))) %>% 
  dplyr::select(IndID, weeks_date, status, sex = Sex,  
                # birth_date, birth_year, last_date, death_year, 
                death_month, 
                season, hs_8km_natal, hs_8km_final,
                nterr_buffer50_first, area_buffer50_first, nterr_dens_buffer50_first,
                nterr_buffer50_last, area_buffer50_last, nterr_dens_buffer50_last)

write.csv(surv_data_git, paste0(procdata_wd, "/data_wolf_survival_table.csv"),
                 row.names = FALSE)
```


# Prepare Reproduction Data

Data used in a parallel approach to survival analysis. Raw data in the monitoring 
data for the wolf territories in Germany. From this, we use:
- Data before year 2021
- Remove single wolves territories (status 3)
- For pair territories (status 2), removed first year 
- Removed territories with 10+ pups (welpen) because we do not have the identities of all reproducing females in territories with multiple reproductions

For pair territories, we need to remove the first year for the reproduction analysis, as they are 
assigned as pair territories retroactively after observing reproduction in the following year and, thus, 
they cannot reproduce in the first year they appear in the database. 

We will analyse the probability of reproduction in a territory and the number of pups per reproduction event 
in relation with habitat suitability, territory density, and experience of breeding female 

**RESPONSE variables**
- Probability of reproduction in a territory
- Number of pups for territories with reproduction 

**EXPLANATORY variables**
- Experience of breeding female (number of years of reproduction), in quadratic form to account for potential optimal time
- Habitat suitability of focal territory (8 km radius of centroid)
- Territory density in radius of 50 km of the focal territory


## Load and prepare reproduction data

```{r load and prepare data}
repro_raw <- read.csv("./data_raw/20220719_Territorien DBBW.csv")

head(repro_raw)
str(repro_raw)
nrow(repro_raw) # [1] 1138

## remove events with 10+ pups
repro_tmp <- repro_raw %>% 
  filter(Mjahr < 2021) %>%
  mutate(Npups = case_when(
    is.na(Welpen) ~ 0, 
    TRUE ~ Welpen)) %>% 
  filter(Npups < 10)
nrow(repro_tmp)

## get pack reproduction  
repro_data_packs <- repro_tmp  %>% 
  filter(Status == 1) %>%  # select only pack data
  filter(Mjahr < 2021)
nrow(repro_data_packs) # 694 rows for pack data

# Get pair reproduction data, remove first year (status = 2)
repro_data_pairs <- repro_tmp %>% 
  filter(Status == 2) %>% 
  filter(Mjahr < 2021) %>% 
  group_by(Territorium, Faehe, Ruede) %>% 
  filter(Mjahr != min(Mjahr))
nrow(repro_data_pairs) # 25 rows for pair reproducing

## Put both datasets together and select only relevant variables
repro_data <- rbind(repro_data_packs, repro_data_pairs)
nrow(repro_data)
# [1] 719

# get total territories
repro_data %>% 
  group_by(Territorium) %>% 
  summarise(n = n()) %>% 
  arrange(-n)
# Max entries = 16 (Years monitored)

length(unique(repro_data$Territorium))
# 205 territories

## retain only relevant variables
repro_data <- repro_data %>% 
  dplyr::select(Territorium, Tkurz, Mjahr, Faehe,  Ruede, Status, Npups, DECLNG, DECLAT, Repro)
```

## Add spatial data  

### Habitat suitability
```{r add habitat suitability data}
nrow(repro_data) # 719

## load habitat suitabilty map (values*1000)
hs_map <- rast("./data_raw/ensemble_samebgpres_weightedmean_projection_3035.tif")
hs_map

## transform reproduction data into spatial data
repro_sf <- st_as_sf(repro_data, coords = c("DECLNG", "DECLAT"), crs = 4326) %>% 
  st_transform(crs = 3035)

## Quick visualization
plot(hs_map)
plot(st_geometry(repro_sf), add = TRUE, col = "red")

## extract habitat suitability values in radius of 8 km of centroid
repro_buffers <- st_buffer(repro_sf, dist = 8000)

repro_hs <- exactextractr::exact_extract(hs_map, repro_buffers, fun = "mean")

wolf_repro_hs <- repro_sf %>%
  mutate(hs_8km = repro_hs / 1000)
  
head(wolf_repro_hs)
nrow(wolf_repro_hs) # 719

summary(wolf_repro_hs)
```

### Territory density

```{r use data from survival extraction}

german_border <- st_read("./data_raw/germany_3035.gpkg", crs = 3035)

wolf_terr_sf
wolf_repro_hs

```

```{r get wolf density values in radius of territories over the years}

## create a loop to extract the territory info for each year of first location
nyears <- length(unique(wolf_repro_hs$Mjahr))
total_area_buffer50 <- pi*50^2
by_year_nterr <- data.frame()

for (i in 1:nyears){
  # select the year
  tmp_year = unique(wolf_repro_hs$Mjahr)[i]
  
  # get individual 50km buffers in year
  poly50_tmp <- wolf_repro_hs %>% 
    filter(Mjahr == tmp_year) %>% 
    st_buffer(dist = 50000) 
  
  # remove the area of the buffers outside germany
  poly50_tmp_crop <- st_intersection(poly50_tmp, german_border)
  
  # add real area for each buffer in km2
  poly50_tmp_crop <- poly50_tmp_crop %>% 
    mutate(area_km2 = as.numeric(
      st_area(poly50_tmp_crop) %>% 
        set_units(km^2)))
  
  # get territory (packs and pairs) polygons in year
  wolf_terr_tmp <- wolf_terr_sf %>% 
    filter(Mjahr == tmp_year) %>% 
    st_buffer(dist = 8000)
  
  # get number of territory polygons intersecting the individual buffer
  intersect50_tmp <- st_intersects(poly50_tmp_crop, wolf_terr_tmp)
  n_int50_tmp <- lengths(intersect50_tmp)
  
  # make into a dataframe
  my_df_tmp <- cbind.data.frame(Territorium = poly50_tmp$Territorium,
                          nterr_50km = n_int50_tmp, 
                          area_buffer50km = poly50_tmp_crop$area_km2,
                          Myear_first = tmp_year) 
  # get territory density using buffer size
  my_df_tmp_dens <- my_df_tmp %>% 
    mutate(densTerr_50km = nterr_50km / (area_buffer50km) * total_area_buffer50) #n territories in equivalent buffer radius

  by_year_nterr <- rbind(by_year_nterr, my_df_tmp_dens)
  }


head(by_year_nterr)
nrow(by_year_nterr) #719
nrow(wolf_repro_hs) #719


## joint n terr to data
wolf_repro_hs_nterr <- wolf_repro_hs %>% 
  left_join(by_year_nterr, by = c("Territorium", "Mjahr" = "Myear_first"))

nrow(wolf_repro_hs_nterr) #719
head(wolf_repro_hs_nterr)
length(unique(wolf_repro_hs_nterr$Territorium )) # 205
```


```{r}
## make dataframe again
wolf_repro_hs_nterr_df <- wolf_repro_hs_nterr %>% 
  mutate(x_coord = st_coordinates(wolf_repro_hs)[,1],
         y_coord = st_coordinates(wolf_repro_hs)[,2]) %>% 
  st_drop_geometry()

head(wolf_repro_hs_nterr_df)
length(unique(wolf_repro_hs_nterr$Territorium)) # 205

```


## Add breeding female experience

```{r add delta breeding ind in territory}
# add breeding female time
wolf_repro_dataset <- wolf_repro_hs_nterr_df %>% 
  group_by(Territorium, Faehe) %>% 
  mutate(N_Years_fem = n(),
         First_year_fem = min(Mjahr)) %>% 
  mutate(Delta_bfem = Mjahr - First_year_fem + 1) %>% 
  as.data.frame()

ggplot(wolf_repro_dataset, aes(x = Delta_bfem, y = Npups)) +
  geom_smooth()
```

check the data and remove unknown females

```{r}
length(unique(wolf_repro_dataset$Faehe)) # 202
length(unique(wolf_repro_dataset$Tkurz))# 205

wolf_repro_dataset %>% 
  filter(is.na(Faehe)) # no obs with Na

wolf_repro_dataset %>% 
  filter(Faehe == "") %>% 
  summarise(n = n()) # 155 unknown

## observations of females without the unknowns
bfem <- wolf_repro_dataset %>% 
  group_by(Faehe) %>% 
  summarise(n = n()) %>% 
  filter(Faehe != "") %>% 
  group_by(n) %>% 
  summarise(nfem = n())
bfem
  
ggplot(bfem, aes(x = n)) +
  geom_bar(aes(y = nfem), stat = "identity")
```
      n  nfem
   <int> <int>
 1     1    66
 2     2    47
 3     3    32
 4     4    16
 5     5    18
 6     6     8
 7     7     9
 8     8     3
 9     9     1
10    10     1

```{r final repro data}
wolf_repro_final <- wolf_repro_dataset %>% 
  filter(Faehe != "")

length(unique(wolf_repro_final$Faehe))
length(unique(wolf_repro_final$Territorium)) 
nrow(wolf_repro_final) 
head(wolf_repro_final)
```

```{r save reproduction data}
## save data for later use 
 # write.csv(wolf_repro_final, paste0(procdata_wd, "/data_wolf_reproduction_envconv.csv"),
 #           row.names = FALSE)
# wolf_repro_final <- read.csv(paste0(procdata_wd, "/data_wolf_reproduction_envconv.csv"))
```

```{r remove sensitive info}
head(wolf_repro_final)

repro_data_git <- wolf_repro_final %>% 
  rename(Territory = Territorium, 
         Tshort = Tkurz , 
         # sYear = Mjahr, 
         Female = Faehe,
         Male = Ruede) %>%  
  group_by(Territory) %>% 
  mutate(TerrID = cur_group_id()) %>% 
  ungroup() %>% 
  group_by(Female) %>% 
  mutate(FemID = cur_group_id()) %>% 
  ungroup() %>% 
  group_by(Male) %>% 
  mutate(MaleID = cur_group_id()) %>% 
  ungroup() %>% 
  mutate(MaleID = case_when(
         Male == "" ~ NaN,
         TRUE ~ MaleID)) %>% 
  dplyr::select(TerrID, FemID, MaleID, Repro, Npups, hs_8km, nterr_50km, area_buffer50km, densTerr_50km, 
                Delta_bfem)

length(unique(repro_data_git$TerrID)) 

write.csv(repro_data_git, paste0(procdata_wd, "/data_wolf_reproduction_table.csv"),
                 row.names = FALSE)
```

