---
title: "explore_data"
author: "Aimara Planillo"
date: "2023-06-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r set workspace}
source("./R/source_packages.R")

procdata_wd <- file.path("./data_proc")
plot_wd <- file.path("./output/figures/")
tmp_wd <- file.path("./output/tmp_output/")
```


# Explore Survival Data

We are going to run a first exploration of the raw data for survival, based on 
visual inspection of sex, age and other explanatory variables

## Load data

```{r load and prepare data}
surv_wolf <- read.csv(paste0(procdata_wd, "/data_wolf_survival_table.csv"))

head(surv_wolf)
nrow(surv_wolf)
# [1] 1054

```


## Get data summaries
```{r summary age class}

## add age class at death
wolf_surv_data <- surv_wolf %>% 
  mutate(age_death = case_when(
    status == 1 ~ "alive",
    status == 0 & weeks_date <= 52 ~"Juvenile",
    status == 0 & weeks_date > 52 & weeks_date <= 104 ~"Subadult",
    status == 0 & weeks_date > 104 ~ "Adult")) %>% 
  mutate(age_death = fct_relevel(age_death,"alive", "Juvenile", "Subadult", "Adult")) 

head(wolf_surv_data)

wolf_surv_data %>% 
  group_by(age_death) %>% 
  summarise(n = n()) 
```

```{r summary explanatory variables}
## by sex
wolf_surv_data %>% 
  group_by(sex) %>% 
  summarise(n = n())
#   sex       n
#   <chr> <int>
# 1 f       480
# 2 m       574

## by season
wolf_surv_data %>% 
  group_by(season) %>% 
  summarise(n = n())
#   season     n
#   <chr>  <int>
# 1 Summer   321
# 2 Winter   733

## mean natal HS
mean(wolf_surv_data$hs_8km_natal)
# [1] 375.1385

## mean final HS
mean(wolf_surv_data$hs_8km_final)
# [1] 338.6975

mean(wolf_surv_data$nterr_dens_buffer50_first)
# [1] 26.87473

mean(wolf_surv_data$nterr_dens_buffer50_last)
# [1] 24.45907

```


## Create summary plots

### All years together
```{r total death by age class}
total_death <- wolf_surv_data %>% 
  filter(age_death != "alive") %>% 
  nrow()
total_alive <- wolf_surv_data %>% 
  filter(age_death == "alive") %>% 
  nrow()

# raw mortality numbers by age
mort_age <- wolf_surv_data %>% 
  group_by(age_death) %>% 
  summarise(total_count = n()) %>% 
  mutate(percentage = total_count/sum(total_count)) %>% 
  ungroup()
  
(plot1 <- ggplot(data = mort_age, aes(x = age_death, y = total_count)) +
    geom_col() +
    geom_text(aes(label = total_count), vjust = -0.5, hjust = 1.5) +
    geom_text(aes(label = paste0("(", round(percentage, 2) * 100, "%)")), 
                  vjust = -0.5, hjust = 0) +
    annotate(geom="text", 
             x = 3.8, y = 400, 
             label=paste0("Total dead = ", total_death),
             color="black") +
    annotate(geom="text", 
             x = 3.8, y = 380, 
             label=paste0("Total alive = ", total_alive),
             color="black") +
    annotate(geom="rect", 
             xmin = 3.3, ymin = 360,
             xmax = 4.3, ymax = 420,
             col = "black", fill = "transparent") +
    ggtitle("Total raw counts per age group") +
    xlab("Age-class at death") +
    ylab("#Individuals") +
    theme_bw()) 

ggsave(plot = plot1, 
       filename = paste0(tmp_wd, "/plot_bars_age_death.png"), 
       dpi = 600, width = 6, height = 5)
```


```{r total death by sex}
# raw mortality numbers by sex
wolf_data_tmp <- wolf_surv_data %>% 
  filter(status == 0) %>% 
  group_by(sex) %>% 
  dplyr::summarise(total_count = n())

(plot2 <- ggplot(data = wolf_data_tmp, aes(x = sex, y = total_count)) +
    geom_col() +
    geom_text(aes(label = total_count), vjust = -0.5) +
    ggtitle("Total raw counts per sex",
            subtitle = "Dead individuals") +
    annotate(geom="text", 
             x = 1, y = 400, 
             label=paste0("Total dead = ", total_death),
             color="black") +
    annotate(geom="text", 
             x = 1, y = 380, 
             label=paste0("Total alive = ", total_alive),
             color="black") +
    annotate(geom="rect", 
             xmin = 0.8, ymin = 370,
             xmax = 1.2, ymax = 410,
             col = "black", fill = "transparent") +
    theme_bw())
```


```{r total death by sex and age}
# raw mortality numbers by age and sex
wolf_data_tmp <- wolf_surv_data %>% 
  group_by(age_death, sex) %>% 
  dplyr::summarise(total_count = n())

(plot3 <- ggplot(data = wolf_data_tmp, aes(x = age_death, y = total_count)) +
  geom_col(aes(fill = sex), position = "dodge") +
    scale_fill_manual(values = c("purple", "orange")) +
    
    geom_text(aes(label = total_count, group = sex),
              position = position_dodge(width = 0.9), vjust = -0.5) +
    ggtitle("Total counts by age class and sex") +
    xlab("Age-class at death") +
    ylab("#Individuals") +
    annotate(geom="text", 
             x = 3.9, y = 250, 
             label=paste0("Total dead = ", total_death),
             color="black") +
    annotate(geom="text", 
             x = 3.9, y = 240, 
             label=paste0("Total alive = ", total_alive),
             color="black") +
    annotate(geom="rect", 
             xmin = 3.3, ymin = 230,
             xmax = 4.5, ymax = 260,
             col = "black", fill = "transparent") +
    theme_bw() +
    theme(plot.title = element_text(face = "bold", size = 16, colour = "black"),
      axis.title = element_text(face = "bold", size = 14, colour = "black"),
          axis.text = element_text(size = 12, colour = "black"), 
          legend.title = element_text(face = "bold", size = 14, colour = "black"), 
          legend.text = element_text(size = 12, colour = "black")) 
          ) 

ggsave(plot = plot3, 
       filename = paste0(tmp_wd, "/plot_bars_age_sex_death.png"), 
       dpi = 600, width = 6, height = 5)  
```

```{r plot percentages of dead by age and sex}
head(wolf_surv_data)

wolf_surv_data %>% 
  filter(status == 0) %>% 
  group_by(sex) %>% 
  summarise(n = n(),
            prop = n() / nrow(surv_wolf[surv_wolf$status == 0,]))
#     sex       n  prop
#   <chr> <int> <dbl>
# 1 f       273 0.461
# 2 m       319 0.539

# percentages of sex within each age class
wolf_data_tmp <- wolf_surv_data %>% 
  group_by(age_death, sex) %>% 
  dplyr::summarise(total_count = n()) %>% 
  ungroup() %>% 
  filter(age_death != "alive") %>%  #remove alive ind
  group_by(age_death) %>% 
  mutate(percentage = total_count/sum(total_count))

(plot4 <- ggplot(data = wolf_data_tmp, aes(x = age_death, y = percentage)) +
  geom_col(aes(fill = sex), alpha = 0.7) +
    geom_text(aes(label = paste0("(",total_count, ")"), group = sex),
              position = position_stack(vjust = 0.5)) +
    geom_text(aes(label = round(percentage, 2), group = sex),
              position = position_stack(vjust = 0.65)) +
    ggtitle("Dead individuals: Percentages of sex by age class") +
    scale_fill_manual(values = c("purple", "orange")) +
    theme_bw() +
        theme(plot.title = element_text(face = "bold", size = 16, colour = "black"),
      axis.title = element_text(face = "bold", size = 14, colour = "black"),
          axis.text = element_text(size = 12, colour = "black"), 
          legend.title = element_text(face = "bold", size = 14, colour = "black"), 
          legend.text = element_text(size = 12, colour = "black")) 
    ) 

ggsave(plot = plot4, 
       filename = paste0(tmp_wd, "/plot_bars_percdeath_age_sex.png"), 
       dpi = 600, width = 6, height = 5)
```


### Along the years 

```{r plot summaries deaths age class year}
head(wolf_surv_data)

# get Oranges palette 5 colors from color brewer
my_palette <- c('#ffffd4','#fed98e','#fe9929','#d95f0e','#993404')

(hm_plot <- wolf_surv_data %>% 
    filter(death_year <= 2020) %>% ## remove few observations after
    filter(age_death != "alive") %>% 
    group_by(death_year, age_death) %>% 
    summarise(n = n()) %>% 
  ggplot(aes(x = death_year, y = age_death, fill = n)) +
  geom_tile(col = "grey90", size = 0.1) +
  scale_fill_gradientn(colors = my_palette, name = "Number of \nindividuals", 
                       limits = c(0,80),
                       breaks = c(1,21,41,61,80)) +
  labs(title = "Wolf deaths by age class along the years") +
  xlab("Monitoring year") +
  ylab("Age class") +
  theme_bw() +
  theme(
    plot.title= element_text(size = 16, face = "bold", colour = "black"), 
    plot.subtitle= element_text(size = 14, face = "italic", colour = "black"), 
    axis.title = element_text(size = 14, face = "bold", colour = "black"),
    axis.text = element_text(size = 14, colour = "black"),
    axis.line = element_line(linewidth = 0.5, colour = "black"),
    plot.margin=unit(c(1,4,1,1),"cm"),
    legend.position = c(1.12, 0.78),
    legend.title = element_text(size = 14, face = "bold", colour = "black"),
    legend.text = element_text(size = 14, colour = "black")
  ))

ggsave(plot = hm_plot, 
       filename = paste0(tmp_wd, "/Plot_heatmap_deaths_age_years.png"), 
       dpi = 600, width = 8, height = 4) 
```



# Explore Reproduction Data

## load data
```{r}
## wolf number of territories
## get territory data and make spatial data
wolf_terr_df <- read.csv("./data_raw/20220719_Territorien DBBW.csv") %>% 
  # remove data after year 2020
  filter(Mjahr < 2021) %>%  
  # remove single territorial individuals
  filter(Status != 3) %>% 
  group_by(Mjahr) %>% 
  summarise(nterr = n())

head(wolf_terr_df)

## wolf reproduction table
wolf_repro <- read.csv(paste0(procdata_wd, "/data_wolf_reproduction_table.csv"))
head(wolf_repro)
```

          
## Plot raw data summaries
```{r plot n reproduction events}
# Summary of reproduction events
summ_table_events <- wolf_repro %>% 
  mutate(pups_fct = case_when(
    Npups == 0 ~ "No",
    Npups > 0 ~ "Yes",
    TRUE ~ "No" # when NA, we assume 0
  )) %>% 
  group_by(sYear , pups_fct) %>% 
  summarise(n = n()) %>% 
  as.data.frame() 

summ_table_events$pups_fct <- fct_relevel(summ_table_events$pups_fct, "Yes", "No")
summ_table_events

# Extract colours from scico
paletteer::paletteer_c(palette = "scico::tofino", n = 7, direction = 1)

my_palette_z <- c("#5D9F57FF", "#6C87CAFF")

ggplot(summ_table_events, aes(x = sYear)) +
  geom_line(aes(y = n, colour = pups_fct), size = 1.5) +
  scale_colour_manual(values = my_palette_z,
                      name = "Reproduction", labels = c("Yes", "No")) +
  ylab("Number of territories") +
  xlab("Year") +
  ggtitle("Reproduction events in wolf territories", 
          subtitle = "Number of events over the years") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12, colour = "black"),
    axis.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.text = element_text(size = 12, colour = "black"),
    plot.title =  element_text(size = 16, colour = "black", face = "bold")
  )
```

```{r plot weighted mean proportion of territories with reproduction}
summ_table_events

prop_events <- summ_table_events %>% 
  group_by(sYear, pups_fct) %>% 
  summarise(n = sum(n)) %>% 
  mutate(prop = n/sum(n), 
         total_year = sum(n)) %>% 
  ungroup() %>% 
  mutate(pups_fct = pups_fct %>% 
           fct_relevel("No","Yes"))

prop_events
wolf_terr_df
# very few territories in teh beginning
## compute average number of territories with reproduction 
## keeping all territories 

prop_pups <- prop_events %>% 
  # filter(total_year > 10) %>% 
  filter(pups_fct == "Yes") 

avg_prop <- weighted.mean(x = prop_pups$prop, w = prop_pups$total_year)
avg_prop
# [1] 0.891844

weighted_var <- Hmisc::wtd.var(x = prop_pups$prop, weights = prop_pups$total_year)
sd_prop <- sqrt(weighted_var)
sd_prop
# [1] 0.04491692

# palette
my_lines <- c("avg_prop" = "dashed")

paletteer::paletteer_c(palette = "scico::cork", n = 7, direction = 1)
my_palette_p <- c( "#E6EEECFF", "#448942FF")

# plot ploportions
ggplot(prop_events, aes(x = sYear , y = prop)) +
  geom_bar(aes( fill = pups_fct), col = "transparent", stat = "identity", alpha = 0.7) +
  scale_fill_manual(values = my_palette_p,
                      name = "Reproduction") +
  geom_area(data = wolf_terr_df, aes(x=Mjahr, y = nterr/max(wolf_terr_df$nterr)),
              alpha = 0.2 , size = 1, colour = "darkred", fill = "darkred", linetype = "dotdash") +
  scale_y_continuous(name = "Proportion of territories", limits = c(0,1),
                     sec.axis = sec_axis(~.* max(prop_events$total_year), name = "Total number of territories")) +
  xlab("Year") +
  geom_abline(intercept = avg_prop, slope = 0, linetype = "twodash", 
              colour = "grey40", size = 1) + # this is the line plot
  geom_line(aes(y = 1.5, linetype = "avg_prop"), colour = "grey40") + # this is just to get the legend
  scale_linetype_manual(values=c("twodash"), name = "", label = "Weighted mean")+
  guides(fill = guide_legend(order = 1), 
              linetype = guide_legend(order = 2)) +
  
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(size = 1, colour = "black"),
    axis.line.y.right = element_line(size = 1, colour = "darkred"),
    axis.title.y.right = element_text(size = 14, colour = "darkred", face = "bold"),
    axis.text.y.right = element_text(size = 12, colour = "darkred"),
    axis.text = element_text(size = 12, colour = "black"),
    axis.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.text = element_text(size = 12, colour = "black"),
    plot.title =  element_text(size = 16, colour = "black", face = "bold"),
    legend.margin = margin(-0.5,0,0,0, unit="cm")
  )

ggsave(filename =  paste0(plot_wd, "/Plot_summary_repro_years_nterr.png"), 
       dpi = 600, height = 4.5, width = 8)
```

```{r plot number of pups}
## for all territories 
wolf_repro %>% 
  # filter(Npups > 0) %>% 
  summarise(mean_pups = mean(Npups), 
            sd_pups = sd(Npups)) 
#   mean_pups  sd_pups
# 1  3.72 2.27

## for territories with reproduction
wolf_repro %>% 
  filter(Npups > 0) %>% 
  summarise(mean_pups = mean(Npups), 
            sd_pups = sd(Npups)) 
#   mean_pups  sd_pups
# 1  4.034755 1.95812

## N pups for territories wiht reproduction    
summ_pups_table <- wolf_repro %>% 
  # make all NA zeros
  filter(Npups > 0) %>% 
  mutate(Npups = replace(Npups, is.na(Npups), 0)) %>% 
  group_by(sYear) %>% 
  summarise(n_terr = n(),
            mean_pups = mean(Npups), 
            sd_pups = sd(Npups)) %>% 
  mutate(type = "AllData") %>% 
  as.data.frame() #%>% 

summ_pups_table

(p <- ggplot(data = summ_pups_table, aes(x = sYear, y = mean_pups)) +
  geom_line(colour = "#AC7825FF", size = 2) +
  geom_ribbon(aes(ymin = mean_pups - sd_pups, ymax = mean_pups + sd_pups),
              fill = "#AC7825FF", alpha = 0.2)) +
  xlab("Year") +
  ylab("Average number of pups") +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    # panel.grid = element_blank(),
    axis.line = element_line(size = 1, colour = "black"),
    axis.line.y.right = element_line(size = 1, colour = "darkred"),
    axis.title.y.right = element_text(size = 14, colour = "darkred", face = "bold"),
    axis.text.y.right = element_text(size = 12, colour = "darkred"),
    axis.text = element_text(size = 12, colour = "black"),
    axis.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.text = element_text(size = 12, colour = "black"),
    plot.title =  element_text(size = 16, colour = "black", face = "bold"),
    legend.margin = margin(-0.5,0,0,0, unit="cm")
  )

ggsave(filename =  paste0(plot_wd, "/Plot_summary_repro_avg_pups_years.png"),
       dpi = 600, height = 4, width = 5)
```


## estimate the distribution that adjust to number of pups and the parameters
```{r distribution of number of pups}
## get data only with reproduction (npups > 0)
npups_table <- wolf_repro %>% 
  filter(Npups > 0) 
hist(npups_table$Npups)

fitdistrplus::descdist(npups_table$Npups, discrete = TRUE)

## try poisson
fit_poisson <- fitdist(npups_table$Npups, "pois")
plot(fit_poisson)

fit_poisson$aic
# [1] 2627.652

## try normal
fit_norm <- fitdist(npups_table$Npups, "norm")
plot(fit_norm)

fit_norm$aic
# [1] 2713.052

## poisson is clearly better

hist(npups_table$pups_numbers)
my_pois <- rpois(n = 10000, lambda = mean(npups_table$Npups))
my_norm <- rnorm(n = 10000, mean = mean(npups_table$Npups),sd = sd(npups_table$Npups))

my_pois_df <- as.data.frame(my_pois)
head(my_pois_df)
my_norm_df <- as.data.frame(my_norm)

distributions_data <- cbind(obs = npups_table$Npups, 
                   exp_pois = my_pois,
                   exp_norm = my_norm) %>% 
  as.data.frame()
dist_data_long <- pivot_longer(distributions_data, cols = c("obs", "exp_pois", "exp_norm"))
dist_data_long <- dist_data_long %>% 
  mutate(dist = name) %>% 
  mutate(dist = dist %>% 
           fct_relevel("obs", "exp_pois"))
str(dist_data_long)

## plot densities
ggplot(dist_data_long, aes(x = value, groups = dist, colour = dist, fill = dist)) +
  geom_density(aes(alpha = dist, linetype = dist), bw = 0.42, position = "identity") +
  scale_colour_manual(values = c("grey20", "orange", "navyblue")) +
  scale_fill_manual(values = c("darkgrey", "orange", "navyblue")) +
  scale_linetype_manual(values = c("solid", "solid", "dashed")) +
  scale_alpha_manual(values = c(1, 0.2, 0.1)) +
  xlab("# Pups") +
  ylab("Density") +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    # panel.grid = element_blank(),
    axis.line = element_line(size = 1, colour = "black"),
    axis.line.y.right = element_line(size = 1, colour = "darkred"),
    axis.title.y.right = element_text(size = 14, colour = "darkred", face = "bold"),
    axis.text.y.right = element_text(size = 12, colour = "darkred"),
    axis.text = element_text(size = 12, colour = "black"),
    axis.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.title = element_text(size = 14, colour = "black", face = "bold"),
    legend.text = element_text(size = 12, colour = "black"),
    plot.title =  element_text(size = 16, colour = "black", face = "bold"),
    legend.margin = margin(-0.5,0,0,0, unit="cm")
  )
```
