---
title: "survival_general"
author: "Aimara Planillo"
date: "2023-06-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare workspace}
source("./R/source_packages.R")
source("./R/source_functions.R")
 
procdata_wd <- file.path("./data_proc")
plot_wd <- file.path("./output/figures/")
tmp_wd <- file.path("./output/tmp_output/")
table_wd <- file.path("./output/tables")
```


# Survival Analysis - General population

Using as explanatory variables *sex*, *season*, *habitat suitability* and *territory density*

Week_date: n of weeks ind was alive 

Response Status: 
- Dead = 1
- Alive = Censored = 0


**EXPLANATORY**
Sex: 
- male
- female

Season:
- Summer = May-Oct (when new pups are born and stay with pack)
- Winter = Nov-Apr 

HS
- Natal territory
- Final territory

Territory density:
- Around Natal territory 
- Around Final territory 

### load survival data
```{r}
wolf_surv_df <- read.csv(paste0(procdata_wd, "/data_wolf_survival_table.csv"))

head(wolf_surv_df)

## Summary number of individuals per status
wolf_surv_df %>% 
  group_by(status) %>% 
  summarise(nind = n())
#  status  nind
#    <int> <int>
# 1      0   588
# 2      1   466
```

### get summaries in data
```{r}
## by sex
wolf_surv_df %>% 
  group_by(sex) %>% 
  summarise(n = n())
# sex       n
#   <chr> <int>
# 1 f       480
# 2 m       574


## by season
wolf_surv_df %>% 
  group_by(season) %>% 
  summarise(n = n())
# sex       n
#   <chr> <int>
# 1 Summer   321
# 2 Winter   733

## mean natal HS
mean(wolf_surv_df$hs_8km_natal)
# [1] 0.3756536
sd(wolf_surv_df$hs_8km_natal)
# [1] 0.101517


## mean final HS
mean(wolf_surv_df$hs_8km_final)
# [1] 0.3388906
sd(wolf_surv_df$hs_8km_final)
# [1] 0.1067531

## mean natal nterr
mean(wolf_surv_df$nterr_dens_buffer50_first)
# [1] 21.16594

sd(wolf_surv_df$nterr_dens_buffer50_first)
# [1] 14.75159


## mean natal nterr
mean(wolf_surv_df$nterr_dens_buffer50_last)
# [1] 21.64657

sd(wolf_surv_df$nterr_dens_buffer50_last)
# [1] 15.77476
```

## test for correlations in explanatory variables
```{r get correlations in data} 
## get numeric explanatory variables
var_surv <- wolf_surv_df %>% 
  dplyr::select(hs_8km_natal, hs_8km_final, nterr_dens_buffer50_first, nterr_dens_buffer50_last)

var_cor <- cor(var_surv, use = "pairwise.complete.obs")

# plot correlation
ggcorrplot(var_cor, hc.order = FALSE, type = "lower",
   lab = TRUE, outline.color = "white", 
   lab_size = 6, tl.cex = 12) +
  theme(plot.background = element_rect(colour = "white"), 
        axis.text = element_text(size = 12)) 

# ggsave(paste0(tmp_wd, "/Survival_general_model_corr_variables.png"),
#        dpi = 600, height = 6, width = 6)
```
Territory density in natal area is correlated with that of final area


## all data - univariate and interaction regressions
We check which variables adjust better to the response
We select he variables to keep by exploring the univariate results, and also the 
value of including an interaction sex:season in the bivariate response. 

```{r cox regression univariate}
head(wolf_surv_df) 

# scale the variables
wolf_surv_df_scl <-  wolf_surv_df %>% 
  mutate(nterr_dens_buffer50_first = scale(nterr_dens_buffer50_first), 
         nterr_dens_buffer50_last = scale(nterr_dens_buffer50_last),
         hs_8km_natal = scale(hs_8km_natal),
         hs_8km_final = scale(hs_8km_final))
  
  
## univariate cox regression
## apply all univariate models in a loop 
colnames(wolf_surv_df_scl)

## numerical or two level covariates
covariates_1 <- c("sex", "season", 
                  "hs_8km_natal", "hs_8km_final", 
                  "nterr_dens_buffer50_first", "nterr_dens_buffer50_last")

univ_formulas_1 <- sapply(covariates_1,
                        function(x) as.formula(paste('Surv(weeks_date, status)~', x)))
                        
univ_models_1 <- lapply( univ_formulas_1, function(x){coxph(x, data = wolf_surv_df_scl)})
univ_results_1 <- lapply(univ_models_1,
                       function(x){ 
                         AICc <- AICc(x)
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value, AICc)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value", "AICc")
                          return(res)
                         })
res_1 <- t(as.data.frame(univ_results_1, check.names = FALSE))
res_1 <- as.data.frame(res_1)
res_1
##save table
# write.csv(res_1, paste0(table_wd, "/Table_survival_general_model_univariate_comparison_AIC.csv"))
```

```{r cox regression sex:season interaction}
## interaction models
covariates_int <- c("sex + season", "sex * season")

int_formulas <- sapply(covariates_int,
                        function(x) as.formula(paste('Surv(weeks_date, status)~', x)))
int_models <- lapply(int_formulas, function(x){coxph(x, data = wolf_surv_df_scl)})

int_tab <- aictab(int_models)

# write.csv(int_tab, paste0(table_wd, "/Table_survival_general_model_bivariate_interation_AIC.csv"),
#           row.names = FALSE)
```

From the exploration:
- The interaction is better than the additive effects
- Nterr density last adjusts better than first to the response

The resulting full model is:
- Sex * season + hs_8km_natal + hs_8km_final + nterr_dens_buffer50_last


## Cox all data - multiple regression

We are going to use an AICc model selection approach

```{r cox prepare multivariate formulas}
head(wolf_surv_df)

## model selection for all data
mod_list_general <- as.list(c(
  "full_model" = as.formula(Surv(weeks_date, status) ~ sex * season + hs_8km_natal + hs_8km_final + nterr_dens_buffer50_last),
  "sxs_first_model" = as.formula(Surv(weeks_date, status) ~ sex * season + hs_8km_natal),
  "sxs_last_model" = as.formula(Surv(weeks_date, status) ~ sex * season + hs_8km_final + nterr_dens_buffer50_last),
  "hs_first_model" = as.formula(Surv(weeks_date, status) ~ hs_8km_natal),
  "hs_last_model" = as.formula(Surv(weeks_date, status) ~ hs_8km_final),
  "nterr_last_model" = as.formula(Surv(weeks_date, status) ~ nterr_dens_buffer50_last),
  "season_model" = as.formula(Surv(weeks_date, status) ~ season),
  "sex_model" = as.formula(Surv(weeks_date, status) ~ sex),
  "null_model" = as.formula(Surv(weeks_date, status) ~ 1))
)

# store in text model for later use
form_text <- vector()
for(i in 1:length(mod_list_general)){
  form_text[i] <- Reduce(paste, deparse(mod_list_general[[i]]))
}

mod_gen_form_table <- cbind.data.frame(Modnames = names(mod_list_general), 
                                       Modformula = form_text)
```

```{r run multivariate regression}
# scale values
wolf_surv_df_scl <- wolf_surv_df %>% 
  mutate(nterr_dens_buffer50_first = scale(nterr_dens_buffer50_first), 
         nterr_dens_buffer50_last = scale(nterr_dens_buffer50_last),
         hs_8km_natal = scale(hs_8km_natal),
         hs_8km_final = scale(hs_8km_final))


general_all_models <- lapply( mod_list_general, function(x){coxph(x, data = wolf_surv_df_scl)})

gen_aictab <- aictab(general_all_models, second.ord = TRUE)
gen_aictab
model_selection_aictab <- gen_aictab %>% 
  as.data.frame() %>% 
  left_join(mod_gen_form_table, by = "Modnames")
model_selection_aictab
# write.csv(model_selection_aictab, paste0(table_wd, "/Table_survival_general_model_multivariate_model.csv"), row.names = FALSE)

## best model
best_formula <- as.formula(model_selection_aictab$Modformula[1])
general_cox <- coxph(formula = best_formula, data = wolf_surv_df_scl)
summary(general_cox)

general_cox_coeftable <- broom::tidy(general_cox)
# write.csv(general_cox_coeftable, 
#            paste0(table_wd, "/Table_survival_general_model_BEST_coeff.csv"), row.names = FALSE)

general_cox_HR_table <- broom::tidy(general_cox, exponentiate = TRUE, conf.int = TRUE)
# write.csv(general_cox_HR_table, 
#            paste0(table_wd, "/Table_survival_general_model_BEST_hazratios.csv"), row.names = FALSE)


# diagnostic for the model
ggcoxdiagnostics(general_cox, type = "deviance", 
                 ox.scale = "linear.predictions")

# get median survival times
## for general plot
general_fit <- survfit(general_cox, na.action = "na.rm") 
median_times <- surv_median(general_fit)
median_times
# strata median lower upper
# 1    All     146    106   360
```

```{r plot pretty Hazard Ratios}
## Customized plot to show the HR
tidycox <- broom::tidy(general_cox, exponentiate = TRUE, conf.int = TRUE)
unique(tidycox$term)
tidycox <- tidycox %>% 
  mutate(variables = case_when(
    term == "sexm" ~ "sex[male]",
    term == "seasonWinter" ~ "season[winter]",
    term == "hs_8km_natal" ~ "HS natal",
    term == "hs_8km_final" ~ "HS final",
    term == "nterr_dens_buffer50_last" ~ "TerrDens final",
    term == "sexm:seasonWinter" ~ "sex*season\n[m*w]")) %>% 
  mutate(variables = variables %>% 
           fct_relevel("sex*season\n[m*w]", "TerrDens final"))

str(tidycox)

p_mid <- ggplot(tidycox, aes(y = variables)) +
  # geom_linerange(aes(xmin = conf.low, xmax = conf.high), 
  #                lwd = 3, colour = "grey60", alpha = 0.7) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), 
                 lwd = 1, colour = "grey60",width = 0.2) +
  geom_point(aes(x = estimate), size = 4) +
  geom_vline(xintercept = 1, linetype="dashed", colour = "darkred") +
  coord_cartesian(ylim = c(1,6), xlim = c(0, 2.5))+
  ylab("") +
  xlab("Hazard Ratio") +
  annotate("text", x = 0.9, y = 6.5, hjust = 1,
           label = 'italic("Decreased risk")', parse = TRUE,
           size = 5, colour = "darkgreen") +
  annotate("text", x = 1.1, y = 6.5, hjust = 0,
           label = 'italic("Increased risk")', parse = TRUE,
           size = 5, colour = "orangered") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", colour = "white"),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.line.x = element_line(colour = "black"),
    # axis.line = element_line(colour = "black"),
    axis.text = element_text(colour = "black", size = 12), 
    axis.title = element_text(colour = "black", size = 14, face = "bold"),
    panel.border = element_blank(),
    )
    

p_left <- ggplot(tidycox, aes(y = variables)) +
  geom_text(aes(x = 0, label = variables), hjust = 0, fontface = "bold", size = 5) +
  geom_text(aes(x = 1.5, label = paste0(round(estimate, 2), 
                                      " (", round(conf.low, 2),
                                      "-", round(conf.high, 2),
                                      ")"),
            hjust = 0, fontface = ifelse(p.value < 0.05, "bold", "plain")), 
            size = 4) +
  theme_void() +
  coord_cartesian(xlim = c(0, 4))
p_left  

p_right <- ggplot(tidycox, aes()) +
  geom_text(aes(x = 0, y = variables, label = round(p.value, 3)),
    hjust = 0,
    fontface = ifelse(tidycox$p.value < 0.05, "bold", "plain"),
    size = 5
  ) +
  annotate("text", label = 'italic("p-value")', x = 0, y = 6.5, parse = TRUE, hjust = 0, size = 5) +
  theme_void()

p_right

# final plot arrangement (patchwork)
layout <- c(
  area(t = 0, l = 0, b = 30, r = 15), # left plot, starts at the top of the page (5) and goes 30 units down and 15 units to the right
  area(t = 0, l = 12, b = 30, r = 25), # middle plot starts a little lower (t=1) because there's no title. starts 1 unit right of the left plot (l=4, whereas left plot is r=3), goes to the bottom of the page (30 units), and 6 units further over from the left plot (r=9 whereas left plot is r=3)
  area(t = 0, l = 24, b = 30, r = 28) # right most plot starts at top of page, begins where middle plot ends (l=9, and middle plot is r=9), goes to bottom of page (b=30), and extends two units wide (r=11)
)
# final plot arrangement
my_HR_plot <- p_left + p_mid + p_right + plot_layout(design = layout)

my_HR_plot

# ggsave(plot = my_HR_plot, 
#        paste0(plot_wd, "/Plot_survival_general_model_HR_pretty.png"),
#        dpi = 600, width = 9, height = 6)
```

```{r plot survival curve - general model}

mydata <- broom::tidy(general_fit)
nrow(mydata) 

## correct the last value for hte confidence interval
mydata$conf.high[229]  <- mydata$conf.high[228]
mydata$conf.low[229]  <- mydata$conf.low[228]

myplot <- ggplot(mydata, aes(time, estimate)) +
  # geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .5, fill = 'blue') +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .5, 
              fill = 'deepskyblue4', colour = "transparent", linetype = 3) +
  geom_step(col = "deepskyblue4", size = 1) +
  # ggtitle("Survival Curve Wolf Germany") +
  # subtitle = "Known birth data") +
  xlab("Weeks") +
  ylab("Survival %") +
  # scale_color_manual(values="#CC6666") +
  geom_segment(aes(x = median_times$median, xend = median_times$median, 
                   y = 0, yend = 0.5), linetype = 6, linewidth = 0.5,  col = "deepskyblue4") +
  geom_segment(aes(x = 0, xend = median_times$median, 
                   y = 0.5, yend = 0.5), linetype = 6, linewidth = 0.5,  col = "deepskyblue4") +
  annotate(
    "text", label = paste0("Median survival time\n", median_times$median, " weeks"),
    x = median_times$median + 10, y = 0, size = 5, colour = "deepskyblue4", 
    fontface = "italic",  hjust = 0, vjust = 0) +
   annotate(
     "text", label = "Model: \nsex * season + HS natal + HS final + TerrDens final",
     x = 230, y = 0.95, size = 5, colour = "grey20", hjust = 0, vjust = 0.5, fontface = 3) +
  theme_bw()+
  theme(
   panel.border = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, colour = "black", face = "bold"),
    axis.text = element_text(size = 14, colour = "black"),
    plot.title = element_text(size = 16, colour = "black", face = "bold"),
    plot.subtitle = element_text(size = 14, colour = "black", face = "italic")
    
  )
myplot


ggsave(plot = myplot,
  paste0(plot_wd, "/Plot_survival_general_model_survival_curve2.png"),
       dpi = 600, height = 6, width = 8)

```

## Get annual survival of all individuals together
```{r}

head(wolf_surv_df)
## divide data into annual observations
wolf_surv_df %>% 
  group_by(status) %>% 
  summarise(nind = n())
#   status  nind
#    <dbl> <int>
# 1      0   588
# 2      1    466

## make a loop to divide each individual into as many rows as years
ind_split <- base::split(wolf_surv_df,
                           wolf_surv_df$IndID)
length(ind_split)
# [1] 1054 ind

new_list <- list()
for(i in 1:length(ind_split)){
  ind_tmp <- ind_split[[i]]
  nyears_tmp <- ind_tmp$weeks_date/52
  ind_tmp$nyears <- ceiling(ind_tmp$weeks_date/52)
  ind_duprows_tmp <- ind_tmp %>% 
  slice(rep(1:n(),each = nyears))
  # make status in every row censored except the last one that we keep as it was
  ind_final <- ind_duprows_tmp %>% 
  mutate(status_ind = status) %>% 
  mutate(status_ind = replace(status_ind, 1:nrow(ind_duprows_tmp)-1, 0)) %>% 
  mutate(weeks_ind = weeks_date) %>% 
  mutate(weeks_ind = replace(weeks_ind, 1:nrow(ind_duprows_tmp)-1, 
                                 values = 52)) %>% 
  mutate(weeks_ind = replace(weeks_ind, nrow(ind_duprows_tmp), 
                                 values = weeks_date - ((nrow(ind_duprows_tmp)-1) * 52)))
  # add to a new list
  new_list[[i]] <- ind_final 
}

head(new_list)
# make the new list into a dataframe again
pop_times_peryear <- do.call(rbind, new_list)

head(pop_times_peryear)

length(unique(pop_times_peryear$IndID))
# [1] 1054 # number of individuals

pop_times_peryear %>% 
  group_by(status_ind) %>% 
  summarise(nind = n())
# status_ind  nind
#           <dbl> <int>
# 1             0   1588
# 2             1    466 # Actual deaths

head(pop_times_peryear)
range(pop_times_peryear$weeks_ind)

# write.csv(pop_times_peryear, paste0(tmp_wd, "/data_population_times_peryear.csv"), row.names = FALSE)
```

## check variable correlation
```{r}
num_vars <- pop_times_peryear %>% 
  dplyr::select(hs_8km_natal , hs_8km_final ,
                nterr_dens_buffer50_first, nterr_dens_buffer50_last )

cor(num_vars)
```
Territory density in natal and final areas is correlated, select the best one for the models

## Univariate and bivariate regressions 

```{prepare variables for univariate models}
# covariate list for all univariate models
head(pop_times_peryear)

## num and two level covariates
my_cov_num <- c("sex", "season", 
                "hs_8km_natal","hs_8km_final",
                "nterr_dens_buffer50_first", "nterr_dens_buffer50_last")

## interaction covariates
my_cov_int <- c("sex + season", "sex * season")
```

#### univ 
```{r pop univariate models}
## Now we use the models with Individual as clustering variables to account for the repetitions over the years 

## models for numerical or two level covariates
univ_formulas_1 <- sapply(my_cov_num,
                        function(x) as.formula(paste('Surv(weeks_ind , status_ind)~', x)))
univ_models_1 <- lapply( univ_formulas_1, function(x){coxph(x, data = pop_times_peryear, cluster = IndID)})

# extract results from the models
univ_results_1 <- lapply(univ_models_1, FUN = get.model.output.cont)
res_pop <- get.univ.table.cont(univ_results_1)
# write.csv(res_pop, paste0(table_wd, "/Table_survival_allpopulation_model_univariate_comparison.csv"))

## models for multilevel covariates
univ_formulas_2 <- sapply(my_cov_int,
                        function(x) as.formula(paste('Surv(weeks_ind, status_ind)~', x)))
univ_models_2 <- lapply(univ_formulas_2, function(x){coxph(x, data = pop_times_peryear,  cluster = IndID)})

res_pop_int <- aictab(univ_models_2)
# write.csv(res_pop_int, paste0(table_wd, "/Table_survival_allpopulation_model_bivariate_interaction_AIC.csv"))
```
Territory density in the final territory adjusts better to the data
The interaction sex * season *improves* the model for population

## multiple regressions 

```{r prepare model selection for multivariate models}
head(pop_times_peryear)

mod_list_pop <- c(
    ## models with sex and season as additive
  "full_model" = as.formula(Surv(weeks_ind, status_ind) ~ sex * season + hs_8km_natal + hs_8km_final + nterr_dens_buffer50_last),
  "sxs_hs_first_model" = (Surv(weeks_ind, status_ind) ~ sex * season + hs_8km_natal),
  "sxs_hs_last_model" = (Surv(weeks_ind, status_ind) ~ sex * season + hs_8km_final),
  "sxs_nterr_model" = (Surv(weeks_ind, status_ind) ~ sex * season + nterr_dens_buffer50_last),
  "hs_model" = as.formula(Surv(weeks_ind, status_ind) ~ hs_8km_natal + hs_8km_final),
  "hs_first_model" = as.formula(Surv(weeks_ind, status_ind) ~ hs_8km_natal),
  "hs_last_model" = as.formula(Surv(weeks_ind, status_ind) ~ hs_8km_final),
  "nterr_last_model" = as.formula(Surv(weeks_ind, status_ind) ~ nterr_dens_buffer50_last), 
  "season_model" = as.formula(Surv(weeks_ind, status_ind) ~  season),
  "sex_model" = as.formula(Surv(weeks_ind, status_ind) ~  sex),
  "null_model" = as.formula(Surv(weeks_ind, status_ind) ~ 1))

form_text <- vector()
for(i in 1:length(mod_list_pop)){
  form_text[i] <- Reduce(paste, deparse(mod_list_pop[[i]]))
}

mod_pop_form_table <- cbind.data.frame(Modnames = names(mod_list_pop), 
                                       Modformula = form_text)
```

### multiv population  
```{r yearly survival pop}
head(pop_times_peryear) 

## Multiple regression with cluster effects for Individual to account for the repetition
pop_all_models <- lapply( mod_list_pop, function(x){coxph(x, cluster = IndID, 
                                                                data = pop_times_peryear, 
                                                                )})
pop_aictab <- aictab(pop_all_models, second.ord = TRUE)
pop_aictab
pop_aictab <- pop_aictab %>% 
  as.data.frame() %>% 
  left_join(mod_pop_form_table, by = "Modnames")
# write.csv(pop_aictab, paste0(table_wd, "/Table_survival_allpopulation_multivariate_modelsel.csv"), row.names = FALSE)


## best model is the full model
best_formula_pop  <- as.formula(pop_aictab$Modformula[1])
pop_cox <- coxph(formula = best_formula_pop, data = pop_times_peryear)
summary(pop_cox)

pop_coeftable <- tidy(pop_cox)
# write.csv(pop_coeftable, paste0(table_wd, "/Table_survival_allpopulation_model_BEST_coef.csv"), row.names = FALSE)

pop_hrtable <- tidy(pop_cox, exponentiate = TRUE, conf.int = TRUE)
# write.csv(pop_hrtable, paste0(table_wd, "/Table_survival_allpopulation_model_BEST_hazratios.csv"), row.names = FALSE)

pop_fit <- survfit(pop_cox, na.action = "na.rm") 
summary(pop_fit)
# get survival and confidence interval
round(pop_fit$surv[length(pop_fit$surv)], 2)
# [1] 0.81
round(pop_fit$lower[length(pop_fit$surv)], 2)
# [1] 0.76
round(pop_fit$upper[length(pop_fit$surv)], 2)
# [1] 0.85

## std error
se <-  round(pop_fit$std.err[length(pop_fit$surv)], 3)
se
# [1] 0.028

# diagnostic for the model
ggcoxdiagnostics(pop_cox, type = "deviance", 
                 ox.scale = "linear.predictions")

## Plot hazard ratios
Pop_HR <- ggforest(pop_cox, data = pop_times_peryear,
                   main = "Hazard ratio - adults known birth") 
Pop_HR
  
```


```{r yearly HR plots manuscript}
## Extract data for forest plot (using own function)
toplot_pop <- broom::tidy(pop_cox, exponentiate = TRUE, conf.int = TRUE)

toplot_pop$model <- "Population"

# rename variables and add significant variable for plotting
toplot_pop_rn <- toplot_pop %>% 
  # rename variables 
   mutate(Variable_name = case_when(
    term == "sexm" ~ "sex[male]",
    term == "seasonWinter" ~ "season[winter]",
    term == "hs_8km_natal" ~ "HS Natal",
    term == "hs_8km_final" ~ "HS Final",
    term == "nterr_dens_buffer50_last" ~ "TerrDens Final",
    term == "sexm:seasonWinter" ~ "sex*season[m*w]")) %>%
  mutate(Variable_name = Variable_name %>% 
           fct_relevel("sex*season[m*w]", "TerrDens Final")) %>% 
  mutate(significant = case_when(
    p.value < 0.05 ~ "Yes",
    TRUE ~ "No")) %>% 
  mutate(estimate2 = round(estimate, 2),
         conflow2 = round(conf.low, 2),
         confhigh2 = round(conf.high, 2))


# Forest Plot
str(toplot_pop_rn)

zp1 <- ggplot(toplot_pop_rn, aes(group = as.factor(model), 
  colour = as.factor(model))) + 
  # future vertical line at 1
  geom_hline(yintercept = 1, colour = "darkred", alpha = 0.5, lty = 2) +
  # line of 95 CI
  geom_linerange(aes(x = Variable_name, ymin = conflow2,
                                ymax = confhigh2), 
                            lwd = 1.5, position = position_dodge(width = -0.8/2)) +
  # # this is the coefficient. No so important the values of ymin and ymax
  geom_pointrange(aes(x = Variable_name, y = estimate2, ymin = conflow2,
                                 ymax = confhigh2, fill = significant),
                             lwd = 1/2, shape = 21, position = position_dodge(width = -0.8/2)) +
  # make it vertical
  xlab("") +
  ylab("Hazard Ratio") +
  coord_flip() + 
  # color for significant
  scale_fill_manual(values = c("White", "black"), 
                    name = "Significant")+
  scale_color_manual(values = scico(4, begin = 0.15, palette = "tokyo"), 
                     name = "Age class") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_blank(), 
    axis.line = element_line(colour = "black"),
    legend.background = element_blank(),
    axis.title = element_text(colour = "black", size = 14, face = "bold"),
    axis.text = element_text(colour = "black", size = 14),
    legend.title = element_text(colour = "black", size = 12, face = "bold"),
    legend.text = element_text(colour = "black", size = 12)) 

zp1

# save plot
# ggsave(plot = zp1, paste0(plot_wd, "/Plot_survival_allpopulation_HR_pretty_scl.png"),
#        dpi = 600, height = 6, width = 8)
```
