---
title: "reproduction"
author: "Aimara Planillo"
date: "2023-06-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare workspace}
source("./R/source_packages.R")
 
procdata_wd <- file.path("./data_proc")
plot_wd <- file.path("./output/figures/")
tmp_wd <- file.path("./output/tmp_output/")
table_wd <- file.path("./output/tables")
```

# Reproduction analysis


**RESPONSE**
- Proportion of territories with reproduction / Probability of reproduction in a territory
- Number of pups for territories with reproduction 

**EXPLANATORY**
- Experience of breeding female (years of reproduction)
- Habitat suitability (in 8 km radius of centroid of territory)
- Wolf density in 50km radius (assuming mirror situation in Poland)

As the territories were not all followed for the same number of years, we are going to use the number of years that 
a territory was monitored as a weighting variable in the analysis

## load and prepare reproduction data
```{r load and prepare data}
repro_tmp <- read.csv(paste0(procdata_wd, "/data_wolf_reproduction_table.csv"))

head(repro_tmp)
str(repro_tmp)
nrow(repro_tmp)

repro_tmp %>% 
  group_by(TerrID) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

length(unique(repro_tmp$TerrID)) # 165
```


```{r}
# add number of years for each territory
nyears2 <- repro_tmp %>% 
  group_by(TerrID) %>% 
  summarise(Terr_years = n()) %>% 
  arrange(-Terr_years)

wolf_repro <- repro_tmp %>% 
  left_join(nyears2, by = "TerrID")

head(wolf_repro)
```

Summaries data
```{r}
# N territories
wolf_repro %>% 
  group_by(TerrID) %>% 
  summarise(n = n()) %>% 
  summarise(n = n())

# N different breeding females
wolf_repro %>% 
  group_by(FemID) %>% 
  summarise(n = n()) %>% 
  summarise(n = n())

# n years per territory
wolf_repro %>% 
  group_by(TerrID) %>% 
  summarise(n = n()) %>% 
  group_by(n) %>% 
  summarise(nterr = n())

# n years per female
wolf_repro %>% 
  group_by(FemID) %>% 
  summarise(n = n()) %>% 
  group_by(n) %>% 
  summarise(nfem = n())

# n pups per reproduction
wolf_repro %>% 
  filter(Repro == 1) %>% 
  summarise(npups = mean(Npups), 
            sdpups = sd(Npups))  
```

After removing unknown females: 
165 territories, 201 known females, maximum 16 years for one territory, 10 years for one female

     npups   sdpups
1 4.075728 2.043328

## Explore variable correlation  

```{r variable correlation}
## check correlations
cor_variables <- wolf_repro %>% 
  dplyr::select(FemExp = Delta_bfem, HS = hs_8km,
                TerrDens = densTerr_50km, Terr_years)

cor_table <- cor(cor_variables)
# plot correlation
ggcorrplot(cor_table, hc.order = FALSE, type = "lower",
   lab = TRUE, outline.color = "white", 
   lab_size = 6, tl.cex = 13) +
  theme(plot.background = element_rect(colour = "white"), 
        legend.title = element_text(size = 14)) 

ggsave(paste0(tmp_wd, "/Repro_variables_correlations_20230912.png"),
       dpi = 600, height = 6, width = 6)
```
No strong correlation among the variables


## model selection list
```{r model list}
## Model selection list, similar to survival models
head(wolf_repro)

## for number of pups
mod_list_npups <- c(
  "full_model" = as.formula(Npups ~ hs_8km * densTerr_50km   + 
                              hs_8km  * (Delta_bfem  + I(Delta_bfem^2)) + 
                              (1|TerrID)), 
  "additive_model" = as.formula(Npups ~ hs_8km + densTerr_50km  + 
                                  Delta_bfem  + I(Delta_bfem^2) + (1|TerrID)),
  "hs_model" = as.formula(Npups ~ hs_8km  + (1|TerrID)),
  "density_model" = as.formula(Npups ~ densTerr_50km  + (1|TerrID )),
  "agefem_quad_model" = as.formula(Npups ~ Delta_bfem + I(Delta_bfem^2) + (1|TerrID)),
  "agefem_model" = as.formula(Npups ~ Delta_bfem + (1|TerrID)),
  "null_model" = as.formula(Npups ~ 1 + (1|TerrID))
)

# store in text model for later use
form_text <- vector()
for(i in 1:length(mod_list_npups)){
  form_text[i] <- Reduce(paste, deparse(mod_list_npups[[i]]))
}
mod_npups_form_table <- cbind.data.frame(Modnames = names(mod_list_npups), 
                                       Modformula = form_text)

## Reproduction probability
mod_list_parepro <- c(
  "full_model" = as.formula(Repro_event ~ hs_8km  * densTerr_50km  + 
                              hs_8km  * (Delta_bfem  + I(Delta_bfem^2)) + (1|TerrID)), 
  "additive_model" = as.formula(Repro_event ~ hs_8km  + densTerr_50km  +
                                  Delta_bfem  + I(Delta_bfem^2) + (1|TerrID)),
  "hs_model" = as.formula(Repro_event ~ hs_8km  + (1|TerrID)),
  "density_model" = as.formula(Repro_event ~ densTerr_50km  + (1|TerrID)),
  "agefem_quad_model" = as.formula(Repro_event ~ Delta_bfem + I(Delta_bfem^2) + (1|TerrID)),
  "agefem_model" = as.formula(Repro_event ~ Delta_bfem + (1|TerrID)),
  "null_model" = as.formula(Repro_event ~ 1 + (1|TerrID))
)

# store in text model for later use
form_text <- vector()
for(i in 1:length(mod_list_parepro)){
  form_text[i] <- Reduce(paste, deparse(mod_list_parepro[[i]]))
}

mod_parepro_form_table <- cbind.data.frame(Modnames = names(mod_list_parepro), 
                                       Modformula = form_text)
```

## Models for prob reproduction
```{r Models for pa reproduction}
## Analyse presence/absence reproduction
##scale data
wolf_repro_pa_scl <- wolf_repro %>% 
  mutate(Repro_event = case_when(
    Repro == 1 ~ 1,
    TRUE ~ 0)) %>% 
  mutate(hs_8km = scale(hs_8km),
         densTerr_50km = scale(densTerr_50km),
         Delta_bfem = scale(Delta_bfem)) %>% 
  ## add number of successes per territory as weights due to binomial distribution
  group_by(TerrID) %>% 
  mutate(terr_weights = sum(Repro_event)) %>% 
  ungroup()

## run models
parepro_all_models <- lapply(mod_list_parepro, function(x){glmer(x, data = wolf_repro_pa_scl, 
                                                        family = "binomial", 
                                                        control = glmerControl(optimizer="bobyqa"),
                                                        weights = Terr_years/max(Terr_years))})

summary(parepro_all_models$density_model)

parepro_aictab <- aictab(parepro_all_models, second.ord = TRUE)
parepro_aictab

parepro_aictab <- parepro_aictab %>% 
  as.data.frame() %>% 
  left_join(mod_parepro_form_table, by = "Modnames")
write.csv(parepro_aictab, paste0(table_wd, "/Table_repro_prob_modsel_weights_20231010.csv"), row.names = FALSE)

## best model: Density model but very close to NULL MODEL!!
parepro_aictab[1:2,]
```

Best model is the density model 
Second best is the null female


## get base probability of reproduction from the null model
```{r null model and backtransform coefficients}
wolf_repro_dataset_pa <-  wolf_repro %>% 
  mutate(Repro_event = case_when(
    Repro == 1 ~ 1,
    TRUE ~ 0)) %>% 
  group_by(TerrID) %>% 
  mutate(Terr_years = n(), 
         terr_weights = sum(Repro_event)) %>% 
  ungroup()

parepro_aictab[1,]

null_repro <- glmer(Repro_event ~ 1 + (1 | TerrID),
            data = wolf_repro_dataset_pa,
            # control = glmerControl(optimizer="bobyqa"),
            family = "binomial",
            weights = Terr_years/max(Terr_years))
summary(null_repro)

saveRDS(null_repro, paste0(tmp_wd, "/Repro_prob_nullModel_model_weights_20231010.rds"))

## get model coefficients
coefs_repromodel <- broom.mixed::tidy(null_repro)
write.csv(coefs_repromodel, paste0(table_wd, "/Table_repro_prob_nullModel_coeff_weights_20231010.csv"), 
          row.names = FALSE)


## backtransform intercept to probabilities
int_coef <- null_repro@beta

odds <- exp(int_coef)
prob <- odds / (1+odds)
prob
# 0.8997892 
```
Intercept: base probability of reproduction a year: 0.90


```{r best model - terr density}
## best model: !!
parepro_aictab[1,]

best_repro_model <- glmer(Repro_event ~ scale(densTerr_50km) + (1 | TerrID),
            data = wolf_repro_dataset_pa,
             control = glmerControl(optimizer="bobyqa"),
            family = "binomial",
            weights = Terr_years/max(Terr_years))

summary(best_repro_model)
plot(ggeffect(best_repro_model))
```

Althought probability of reproduction seems to increase with density, not significant effect

```{r Plot model pa reproduction}
pdt <- ggeffect(best_repro_model, terms = "densTerr_50km") %>% 
ggplot(aes(x = x, y = predicted)) +
  geom_line(col = "brown", linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              fill = "grey40", alpha = 0.25) +
  ylab("Probability of Reproduction") +
  coord_cartesian(ylim = c(0.70, 1)) +
  xlab("Territory density") +
  ggtitle("") +
    theme_bw() +
  theme(
    panel.border = element_blank(),
    axis.line = element_line("black"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12)
  )

ggsave(plot = pdt,
       filename = paste0(plot_wd, "/Plot_prob_model_effects_density_weights_20231010.png"),
       dpi = 600, height = 4, width = 4)
```



## models for N pups

```{r Models N pups}
## only for those cases with pups (removing the zeros)
##scale data
wolf_repro_scl <- wolf_repro %>% 
  filter(!is.na(Npups)) %>% 
  filter(Npups > 0) %>% 
  mutate(hs_8km = scale(hs_8km)[,1],
         densTerr_50km = scale(densTerr_50km)[,1],
         Delta_bfem = scale(Delta_bfem)[,1]) %>% 
  group_by(TerrID) %>% 
  mutate(terr_weights = n()) %>% 
  ungroup()

## run models
npup_all_models <- lapply(mod_list_npups, function(x){glmer(x, data = wolf_repro_scl, 
                                                        family = "poisson",
                                                        weights = Terr_years )})

npup_aictab <- aictab(npup_all_models, second.ord = TRUE)
npup_aictab

npup_aictab <- npup_aictab %>% 
  as.data.frame() %>% 
  left_join(mod_npups_form_table, by = "Modnames")
write.csv(npup_aictab, paste0(table_wd, "/Table_repro_npups_modsel_weights_20231010.csv"), row.names = FALSE)

## best model
npup_aictab[1,]
# we run it scaling within the formula to plot with real values later

## select only observations with reproduction
wolf_repro_withpups <- wolf_repro %>% 
  filter(!is.na(Npups)) %>% 
  filter(Npups > 0) %>% 
   group_by(TerrID) %>% 
  mutate(Terr_years = n()) %>% 
  ungroup()


#full model
Npups_model <- glmer(Npups ~ scale(hs_8km) * scale(densTerr_50km) + 
                       scale(hs_8km) * (scale(Delta_bfem) + scale(I(Delta_bfem^2))) + 
                       (1|TerrID),
            data = wolf_repro_withpups,
            family = "poisson", 
            weights = Terr_years)

summary(Npups_model)
saveRDS(Npups_model, paste0(tmp_wd, "/Repro_Npups_bestModel_model_weights_20231010.rds"))
## get model coefficients
coefs_pupmodel <- broom.mixed::tidy(Npups_model)
write.csv(coefs_pupmodel, paste0(table_wd, "/Table_repro_npups_model_BEST_coeff_weights_20231010.csv"), 
          row.names = FALSE)
```

```{r Plot model N pups with both interactions}
# linear effects
plot(ggeffect(Npups_model))
# interaction effects
plot(ggeffect(Npups_model, terms = c("densTerr_50km", "hs_8km")))
plot(ggeffect(Npups_model, terms = c("Delta_bfem", "hs_8km")))


phs_dt <- ggeffect(Npups_model, terms = c("densTerr_50km", "hs_8km")) %>% 
ggplot(aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(linewidth = 1, aes (linetype = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), col = "transparent", alpha = 0.25) +
  scale_colour_manual(values = c("palevioletred4", "tan", "seagreen"), name = "Habitat \nSuitability") +
  scale_fill_manual(values = c("palevioletred4", "tan", "seagreen"), name = "Habitat \nSuitability") +
  scale_linetype_manual(values = c(3,2,1),  name = "Habitat \nSuitability") +
  guides(colour = guide_legend(reverse = TRUE),
         fill = guide_legend(reverse = TRUE), 
         linetype = guide_legend(reverse = TRUE)) +
  ylab("Number of pups") +
  coord_cartesian(ylim = c(2, 5.5)) +
  xlab("Territory Density") +
  ggtitle("") +
   theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.background = element_rect(fill = "white", colour = NULL),
    axis.line = element_line("black"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold")
  )
phs_dt


pda_dt <- ggeffect(Npups_model, terms = c("Delta_bfem", "hs_8km")) %>% 
ggplot(aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(linewidth = 1, aes (linetype = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), col = "transparent", alpha = 0.25) +
  scale_colour_manual(values = c("palevioletred4", "tan", "seagreen"), name = "Habitat \nSuitability") +
  scale_fill_manual(values = c("palevioletred4", "tan", "seagreen"), name = "Habitat \nSuitability") +
  scale_linetype_manual(values = c(3,2,1),  name = "Habitat \nSuitability") +
  guides(colour = guide_legend(reverse = TRUE),
         fill = guide_legend(reverse = TRUE), 
         linetype = guide_legend(reverse = TRUE)) +
  ylab("Number of pups") +
  coord_cartesian(ylim = c(2, 6.5)) +
  xlab("Experience of Breeding female") +
  ggtitle("") +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.background = element_rect(fill = "white", colour = NULL),
    axis.line = element_line("black"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold")
  )

pda_dt

newplot <- ggarrange(phs_dt, pda_dt, ncol = 2, 
                     common.legend = TRUE, legend = "right") +
  theme(panel.background = element_rect(fill = "white", colour = NULL)) 

newplot <- (phs_dt + pda_dt) +
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(face = 'bold'), 
                                            legend.position = "right") 
newplot <- newplot + plot_layout(guides = "collect")

ggsave(plot = newplot,
       filename = paste0(plot_wd, "/Plot_npups_model_effects_weights_20231010.png"),
       dpi = 600, height = 4, width = 8)
```

