---
title: "growth_estimation"
author: "A. Planillo"
date: "2024-02-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Estimates of population growth (lambda)

3 age classes Age-structured Matrix Population Model 
Adapted from code provided by Henrik Andren 

```{r}
rm(list=ls())
library(popbio)
```


## Matrix 3 age classes
Leslie matrix with 3 age classes (juvenile (0 to 1 yrs); subadults (1 to 2 yrs); adults (2+))
Post breeding

### define parameters
```{r parameters}
surv_juv <- 0.75
surv_subad <- 0.75
surv_ad <- 0.88

prop_rep <- 0.88
litter_size <- 4.08
p_fem <- 0.5 # proportion of female in the litter

stages<-c("juvenile", "subadults", "adults") 
```

### built population matrix
```{r}
AA <- matrix(c(0, 0, p_fem * prop_rep * litter_size * surv_ad,
               surv_juv, 0, 0,
               0, surv_subad, surv_ad),
             nrow = 3, byrow = TRUE, dimnames = list(stages,c("","","")) )
```

## get population growth rate
```{r}
eigen.analysis(AA)$lambda1
# [1] 1.360259
log(eigen.analysis(AA)$lambda1)
```

## observed population growth
```{r}
nterr_2010 <- 14 
nterr_2020 <- 185 # numbers from the repro table^, removing single individuals

growth <- (nterr_2020/nterr_2010)^1/10
growth
# [1] 1.32
```


## incorporate uncertainty estimates - Stochastic version
```{r}
n <- 10000 # number of iteration
lambda.st <- numeric(n)
sur.juv <- numeric(n)
sur.subad <- numeric(n)
sur.ad <- numeric(n)
prop.rep <- numeric(n)
litter.size <- numeric(n)

mj <- 0.75 # mean juvenile survival
sj <- 0.06 # standard deviation juvenile survival

## Moment matching for beta distribution
ja <- (mj^2-mj^3-mj*sj^2)/sj^2
jb <- (mj-2*mj^2+mj^3-sj^2+mj*sj^2)/sj^2

ms <- 0.75 # mean subadult survival
ss <- 0.10 # standard deviation subadult survival

## Moment matching for beta distribution
ssa <- (ms^2-ms^3-ms*ss^2)/ss^2
ssb <- (ms-2*ms^2+ms^3-ss^2+ms*ss^2)/ss^2

ma <- 0.88 # mean adult survival
sa <- 0.03 # standard deviation adult survival

## Moment matching for beta distribution
aa <- (ma^2-ma^3-ma*sa^2)/sa^2
ab <- (ma-2*ma^2+ma^3-sa^2+ma*sa^2)/sa^2

mr <- 0.88 # mean proportion female reproducing
sr <- 0.05 # standard deviation proportion female reproducing
## Moment matching for beta distribution
ra <- (mr^2-mr^3-mr*sr^2)/sr^2
rb <- (mr-2*mr^2+mr^3-sr^2+mr*sr^2)/sr^2

mp <- 4.08 # mean litter size
sp <- 0.14 # 2.03 standard deviation litter size; SE = 2.03/sqrt(205) = 0.14

## Moment matching for gamma distribution
a.gam <- mp^2/sp^2
b.gam <- mp/sp^2

p.fem <- 0.5 ## p.fem # constant 0.5

stages <- c("juvenile", "subadults", "adults")
```

## Run the simulations
```{r}
for (i in 1:n){
  sur.juv[i] = rbeta(1,ja,jb)
  sur.subad[i] = rbeta(1,ssa,ssb)
  sur.ad[i] = rbeta(1,aa,ab)
  prop.rep[i] = rbeta(1,ra,rb)
  litter.size[i] = rgamma(1,a.gam,b.gam)
  
  AA.st <- matrix(c(
    0,0,p.fem*prop.rep[i]*litter.size[i]*sur.ad[i],
    sur.juv[i],0,0,
    0,sur.subad[i],sur.ad[i]),
    nrow = 3, byrow = TRUE,dimnames=list(stages,c("","","")) )
    lambda.st[i] <- eigen.analysis(AA.st)$lambda1
  
}

mean(lambda.st)
median(lambda.st)
sd(lambda.st)
hist(lambda.st, breaks=100, freq=F)
lines(density(lambda.st), col="red", lwd=2)
abline(v=mean(lambda.st), col="blue", lwd=2)

hist(log(lambda.st), breaks=100, freq=F)
lines(density(log(lambda.st)), col="red", lwd=2)
abline(v=mean(log(lambda.st)), col="blue", lwd=2)

## simulated data
mean(sur.juv)
mean(sur.subad)
mean(sur.ad)
mean(prop.rep)
mean(litter.size)
```


## Plot the density functions for survival and reproduction
simulated data and distributions
```{r}
xx <- seq(0,1,0.001)
hist(sur.juv, freq=F, breaks=100)
p.juv <- dbeta(xx,ja,jb)
lines(typ="l",xx,p.juv ,col="red")

p.sub <- dbeta(xx,ssa,ssb)
hist(sur.subad, freq=F, breaks=100)
lines(typ="l",xx,p.sub ,col="red")

p.ad <- dbeta(xx,aa,ab)
hist(sur.ad, freq=F, breaks=100)
lines(typ="l",xx,p.ad ,col="red")

p.rep <- dbeta(xx,ra,rb)
hist(prop.rep, freq=F, breaks=100)
lines(typ="l",xx,p.rep ,col="red")

x2 <- seq(0,10,0.01)
hist(litter.size, freq=F, breaks=100)
p.lit = dgamma(x2,a.gam,b.gam)
lines(typ="l",x2,p.lit ,col="red")

```

## Sensitivities
```{r}
eigout <- eigen.analysis(AA)

eigout$lambda1
eigout$sensitivities
```


# estimate population growth at low and high habitat suitability

## Low habitat suitability

Estimate extremes of hte habitat suitability effects, taking into accountes changes to:
- Juvenile survival
- Subadult survival
- Npups in reproduction

No changes to adult survival or probability of reproduction, as those were not observed in the analyses

### estimate survivals for low habitat suitability
```{r juvenile survival}
## juveniles (using the dataset from the survival analyses)
juv_cox <- coxph(Surv(weeks_juveniles, status_juveniles) ~ sex + season + hs_8km_natal + nterr_dens_buffer50_first, data = juv_surv_scl)

juveniles_fit <- survfit(juv_cox, na.action = "na.rm") 

# get survival and confidence interval
round(juveniles_fit$surv[length(juveniles_fit$surv)], 2)

## predict hs 
juv_hs_df <- with(juv_surv_scl,
               data.frame(
                 sex = c("f"),
                 season = c("Summer"),
                 hs_8km_natal = c(min(hs_8km_natal),
                                 mean(hs_8km_natal),
                                 max(hs_8km_natal)), 
                  nterr_dens_buffer50_first = mean(nterr_dens_buffer50_first)))
                   
fit <- survfit(juv_cox, newdata = juv_hs_df)
round(fit$surv[nrow(fit$surv),], 2)


surv_juv_low <- 0.58  
surv_juv_high <- 0.85
```

```{r subadult survival}
## subadults (using the dataset from the survival analyses)
subad_cox <- coxph(Surv(weeks_subadults, status_subadults) ~ sex * season + hs_8km_natal + nterr_dens_buffer50_first + hs_8km_final + nterr_dens_buffer50_last, data = sub_surv_scl)

sub_fit <- survfit(subad_cox, na.action = "na.rm") 

# get survival and confidence interval
round(sub_fit$surv[length(sub_fit$surv)], 2)

## predict hs 
sub_hs_df <- with(sub_surv_scl,
               data.frame(
                 sex = c("f"),
                 season = c("Summer"),
                 hs_8km_natal = c(min(hs_8km_natal),
                                 mean(hs_8km_natal),
                                 max(hs_8km_natal)),
                 hs_8km_final = mean(hs_8km_natal),
                 nterr_dens_buffer50_first = mean(nterr_dens_buffer50_first), 
                 nterr_dens_buffer50_last = mean(nterr_dens_buffer50_last)))
                   
sub_fit2 <- survfit(subad_cox, newdata = sub_hs_df)
round(sub_fit2$surv[nrow(sub_fit2$surv),], 2)


surv_sub_low <- 0.66  
surv_sub_high <- 0.81
```


## estimate reproduction effects
```{r}

prop_rep_low <- prop_rep_high <- 0.88 ## this value does not change

## number of pups
Npups_model <- readRDS(paste0(tmp_wd, "/Repro_Npups_bestModel_model_weights_20231010.rds"))
Npups_model@call
exp(mean(predict(Npups_model)))

wolf_repro_withpups <- wolf_repro_withpups %>% 
   filter(!TerrID %in% c(54, 71)) ## remove territories not analysed
  

npups_lowhs_df <- with(wolf_repro_withpups,
               data.frame(
                 TerrID = unique(wolf_repro_withpups$TerrID), 
                 hs_8km = c(min(hs_8km)),
                 densTerr_50km = mean(densTerr_50km), 
                 Delta_bfem = mean(Delta_bfem)))

npups_highhs_df <- with(wolf_repro_withpups,
               data.frame(
                 TerrID = unique(wolf_repro_withpups$TerrID), 
                 hs_8km = c(max(hs_8km)),
                 densTerr_50km = mean(densTerr_50km), 
                 Delta_bfem = mean(Delta_bfem)))


lowhs_npups <- mean(predict(Npups_model, npups_lowhs_df))
high_npups <- mean(predict(Npups_model, npups_highhs_df))

```


```{r}
surv_subad <- 0.75
surv_ad <- 0.88

prop_rep <- 0.88
litter_size_low <- exp(lowhs_npups)
litter_size_high <- exp(high_npups)

p_fem <- 0.5 # proportion of female in the litter

stages<-c("juvenile", "subadults", "adults") 
```


### built population matrices
```{r}
AA_low <- matrix(c(0, 0, p_fem * prop_rep * litter_size_low * surv_ad,
               surv_juv_low, 0, 0,
               0, surv_sub_low, surv_ad),
             nrow = 3, byrow = TRUE, dimnames = list(stages,c("","","")) )


AA_high <- matrix(c(0, 0, p_fem * prop_rep * litter_size_high * surv_ad,
               surv_juv_high, 0, 0,
               0, surv_sub_high, surv_ad),
             nrow = 3, byrow = TRUE, dimnames = list(stages,c("","","")) )
```

## get population growth rate
```{r}
eigen.analysis(AA_low)$lambda1
# [1] 1.221282
eigen.analysis(AA_high)$lambda1
# [1] 1.486122
```
